Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-08-31 final.tex
up

{\bf Home-spaces}
 In 1986, Memmi and Vautherin introduced the notion of home-space~\cite{DBLP:conf/ac/MemmiV86} for a system $\mathscr{S} = (S,\rightarrow )$ with an initial state $s_0$ : a subset $H \subseteq S$ is an \emph{home-space}  if 
$\post^*(s_0) \subseteq \pred^*(H)$. If the home-space contains a single element, this element is an {\em home-state}.
It could be easily generalized for two subsets $X,H$ and we say that $H$ is an \emph{home-space for $X$} if $\post^*(X) \subseteq \pred^*(H)$. In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (only published as an internal report). In 2023, Jancar and Leroux proved the decidability of the (complete) semilinear home-space problem ($X$ and $H$ are both semilinear)  in VASS \cite{DBLP:journals/corr/abs-2207-02697}.
%%%%%

@@

\begin{example}\label{Example}
{(\bf Reset-VASS example).}
Consider the 
 reset-VASS $V_1$ from Figure~\ref{r-V}.
Consider first $\Safe = \{q_0(n) \mid n \in \N\}$; In this case,  {resilience}, 
{$k$-resilience} and {bounded resilience} are not satisfied: there is no  way back from $q_2$ or $q_1$ to $q_0$. Consider now $\Safe = \{q_1(0)\} $; then {resilience} hold: it is possible to reach $q_1(0)$ from any state of $V_1$. However, {bounded resilience} does not hold, since, for all $n\in \N$, from $q_1(n)$, there is no path of length less than $n$ towards $q_1(0)$. Consider now the 
 reset-VASS $V_2$, which is $V_1$ plus an $+1$ transition from $q_1$ back to $q_0$. In $V_2$, {bounded resilience} hold, since $6$-resilience hold. 
Indeed, from any state with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the (unique) reset transition. From $q_1(n)$ or $q_0(n)$ with $n \in \{1, 2, 3, 4, 5, 6\}$ it is possible to reach $q_1(0)$ in $6$ steps or less. From 
$q_1(n)$ or $q_0(n)$ with $n \in \{7,8\}$ it is possible to reach $q_0(10)$ in $3$ steps or less, from which $q_1(0)$ is reachable in $2$ steps. 
\end{example}

@@

\noindent
{\bf On effectivity}
Let a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$. We say that $\mathscr{S}$ is {\em effective} if there exists a pair of algorithms
($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
$M_\leq$ the ordering relation “$\leq$”.
We say that $\mathscr{S}$ is {\em post-effective} if it is effective, and if there
exists an algorithm that computes $|\post(x)| \in \N \cup \{\infty\}$ 
on input $x$, with $x \in X_i $. 
We say that $\mathscr{S}$ has {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
%
We say that $\mathscr{S}$
is {\em ideal-effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a state $s$
to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
%

>>>>

{\bf Home-spaces.}
 In 1986, Memmi and Vautherin introduced the notion of home-space~\cite{DBLP:conf/ac/MemmiV86} for a system $\mathscr{S} = (S,\rightarrow )$ with an initial state $s_0$ : a subset $H \subseteq S$ is an \emph{home-space}  if 
$\post^*(s_0) \subseteq \pred^*(H)$. If the home-space contains a single element, this element is an {\em home-state}.
It could be easily generalized for two subsets $X,H$ and we say that $H$ is an \emph{home-space for $X$} if $\post^*(X) \subseteq \pred^*(H)$. In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (only published as an internal report). In 2023, Jancar and Leroux proved the decidability of the (complete) semilinear home-space problem ($X$ and $H$ are both semilinear)  in VASS \cite{DBLP:journals/corr/abs-2207-02697}.
%%%%%

@@

\begin{example}\label{Example}
{(\bf Reset-VASS example).}
Consider the 
 reset-VASS $V_1$ from Figure~\ref{r-V}.
Consider first $\Safe = \{q_0(n) \mid n \in \N\}$; In this case,  {resilience}, 
{$k$-resilience} and {bounded resilience} are not satisfied: there is no  way back from $q_2$ or $q_1$ to $q_0$. Consider now $\Safe = \{q_1(0)\} $; then {resilience} hold: it is possible to reach $q_1(0)$ from any state of $V_1$. However, {bounded resilience} does not hold, since, for all $n\in \N$, from $q_1(n)$, there is no path of length less than $n$ towards $q_1(0)$. Consider now the 
 reset-VASS $V_2$, which is $V_1$ plus a $+1$ transition from $q_1$ back to $q_0$. In $V_2$, {bounded resilience} hold, since $6$-resilience hold. 
Indeed, from any state with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the (unique) reset transition. From $q_1(n)$ or $q_0(n)$ with $n \in \{1, 2, 3, 4, 5, 6\}$ it is possible to reach $q_1(0)$ in $6$ steps or less. From 
$q_1(n)$ or $q_0(n)$ with $n \in \{7,8\}$ it is possible to reach $q_0(10)$ in $3$ steps or less, from which $q_1(0)$ is reachable in $2$ steps. 
\end{example}

@@

\noindent
{\bf On effectivity.}
Let a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$. We say that $\mathscr{S}$ is {\em effective} if there exists a pair of algorithms
($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
$M_\leq$ the ordering relation “$\leq$”.
We say that $\mathscr{S}$ is {\em post-effective} if it is effective, and if there
exists an algorithm that computes $|\post(x)| \in \N \cup \{\infty\}$ 
on input $x$, with $x \in X_i $. 
We say that $\mathscr{S}$ has {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
%
We say that $\mathscr{S}$
is {\em ideal-effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a state $s$
to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
%
